---
title: "DATA607 - Assignment1"
author: "Rob Welk"
date: "February 2, 2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE)
```
The code in this document demonstrates how to load data from a structured online source and apply basic data transformations.

## **Load packages** 
```{r Load_packages, message=F}
library(tidyverse)
library(plyr)
```


## **Import mushroom data as a tibble**

```{r import_from_source}
import.data <- as_tibble(read.csv("https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data", header = F))
```

## **Select the subset of columns that classifies mushroom cap properties and mushroom habitat with renamed variables**

```{r select_rename, include=TRUE}
dat <- import.data %>% select(c("food.type"="V1","cap.shape"="V2", 
                                "cap.surface"="V3","cap.color"="V4",
                                "habitat"="V23"))
```

## **Revalue factor levels**

```{r revalue_factors_levels}


dat$food.type <- revalue(dat$food.type, c("e"="edible","p"="poisonous"))
dat$cap.shape <- revalue(dat$cap.shape, c("b"="bell","c"="conical","x"="convex",
                                            'f'="flat", 'k'="knobbed",'s'="sunken")) 
dat$cap.surface <- revalue(dat$cap.surface,  c('f'="fibrous",'g'="grooves", 'y'="scaly", 
                                              's'="smooth"))
dat$cap.color <- revalue(dat$cap.color, c('n'='brown', 'b'='buff','c'='cinnamon','g'='gray',
                                          'r'='green', 'p'='pink','u'='purple','e'='red',
                                          'w'='white', 'y'='yellow'))
dat$habitat <- revalue(dat$habitat, c('g'='grasses', 'l'='leaves', 'm'='meadows','p'='paths',
                                      'u'='urban', 'w'='waste','d'='woods'))
```

## **View new tibble with basic transformations**
```{r show_tibble, echo=FALSE}
dat
```

## **Create barplots**
```{r ggplot_barplots}
p1 <- ggplot(data=dat) 

p1 + geom_bar(mapping=aes(x=cap.shape, fill=food.type), position="dodge")
p1 + geom_bar(mapping=aes(x=cap.color, fill=food.type), position="dodge")
p1 + geom_bar(mapping=aes(x=cap.surface, fill=food.type), position="dodge")
p1 + geom_bar(mapping=aes(x=habitat, fill=food.type), position="dodge")
```


# Published to Rpubs on 2/2/19 @ 2:58 pm http://rpubs.com/robertwelk/DATA607_assignment1

https://github.com/robertwelk/DATA607_Assignment1.git