---
title: "DATA607 - Assignment 2"
author: "Rob Welk"
date: "February 10, 2019"
output: html_document
---
This document shows how to retrieve tables from a MySQL database (v.8.0.12) and use relational tables in tidyverse.  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load packages
```{r load-needed-packages, eval=T, message=FALSE}
library(dbConnect)
library(RMySQL)
library(tidyverse)
```


#Establish the conncection with MySQL

permissions for user 'DATA606' were provided in the SQL script on GitHub
```{r connect, eval=T}
mydb = dbConnect(MySQL(), user= "DATA606", host= "127.0.0.1", password= "abc123", dbname= "movie ratings", port=3306)
```

**List the tables in the schmema**   
There are 3 related table in this database.  
'ratings' provides a a rating from 1-5 for each movie given by each reviewer  
'reviewers' provides information on each movie reviewer  
'movie' provides information for each title  

**Find tables available in the schema**  
```{r, eval=T}

dbListTables(mydb)
```

**Find fields within a table**
```{r, eval=TRUE}
dbListFields(mydb, 'ratings') 
```

#Retrieving tables from MySQL
*&Send a query to MySQL for each table and store result as a tibble**
```{r,eval=T}
q1 <- dbSendQuery(mydb, 'select * from ratings')
ratings <- as_tibble(fetch(q1, n=-1))
ratings

q2 <- dbSendQuery(mydb, 'select * from movies')
movies <- as_tibble(fetch(q2,n=-1))
movies

q3 <- dbSendQuery(mydb, 'select * from reviewers')
reviewers <- as_tibble(fetch(q3,n=-1))
reviewers
```

#Using relational tables in tidyverse  

*Which title had the highest average rating?*
```{r}
ratings %>% group_by(MovieID) %>% summarise('avg'=mean(Rating)) %>% left_join(movies, by="MovieID") %>% filter(avg==max(avg)) %>% select(Title)
```

*How many perfect ratings were there? Which movies and who gave the the review?*
```{r}
ratings %>% left_join(movies,by="MovieID") %>% left_join(reviewers,by="ReviewerID") %>% filter(Rating==5) %>% select(Rating,Title,Name) %>% arrange(Title)
```

*There were two documentaries about the failed Fyre Festival - how did ratings vary by reviewer*
```{r, eval=TRUE}
ratings %>% left_join(movies,by="MovieID") %>% left_join(reviewers,by="ReviewerID") %>% 
  filter(Title=="Fyre" | Title=="Fyre Fraud") %>%
  ggplot(aes(x=Name, y=Rating, fill=Title)) + geom_bar(stat="identity", position='dodge')
```


