---
title: "DATA607 - HW#3"
author: "Rob Welk"
date: "February 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Original problem
```{r}
library(stringr)
raw.data <- "555-1239Moe Szyslak (636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"

#code given by Munzert etal. to extract names of characters
names <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))

```

#3.a Rearrange vector so that all names conform to first_name last_name standard.
```{r}
# remove titles from rev. lovejoy and dr. hibbert  
# pattern: titles have periods and are at least 2 letters
names2 <- str_replace(names, "[:alpha:]{2,3}\\.", "")

# switch order for homer and burns
# pattern: in last_name, first_name notation there is a comma always
# detect this pattern in the vector by looping through the character vector and applying conditional logic 
# if it exists, split the string at the comma then switch the order

for(i in 1:length(names)){
  if(str_detect(names2[i],",")){
  spl <- str_split(names2[i],",")
  names2[i] <- str_c(spl[[1]][2]," ",spl[[1]][1])
  }
}
firstname_lastname <- str_trim(names2, side="both")
firstname_lastname
```

#3.b Contsruct a logical vector indicating whether the character has a title.
```{r}
# detect the pattern
title.names <- str_detect(names, "[:alpha:]{2,3}\\.")
# name the vector
names(title.names) <- firstname_lastname
title.names
```

#3.c Construct a logical vector indicating whether a character has a second name.
```{r}
#pattern: a character with a second name will have more than two groups of text when split by blank space
second.name <- sapply(str_split(firstname_lastname, " "), length) > 2
names(second.name) <- firstname_lastname
second.name

```


#4. Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regex

```{r}
#(a) any integer with a dollar sign at the end
pattern.a <- "[0-9]+\\$"
example.a <- "4235$"
str_extract(example.a, pattern.a)

```

```{r}
# (b) a literal b then any lowercase letter (1 to 4 of them) followed by another b 
pattern.b <- "\\b[a-z]{1,4}\\b"
example.b <- "barb"
str_extract(example.b, pattern.b)
```

```{r}
#(c) any character 0 or more times then a literal dot then 'txt'
pattern.c <-  ".*\\.txt$"
example.c <- "anything.txt"
str_extract(example.c, pattern.c)
```
```{r}

#(d) a two digit number /  a two digit number / a four digit number
pattern.d <- "\\d{2}/\\d{2}/\\d{4}"
example.d <- "07/14/1984"
str_extract(example.d, pattern.d)
```