---
title: "DATA607 - Project 1"
author: "Rob Welk"
date: "February 24, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo =TRUE)
```

#Problem Statement
A document contianing information from a chess tournament is provided as an unstructured text file. The code below attempts to tidy the raw string data into a normalized table that is exported as a comma separated value file.  The information to extract for each chess player includes name, state, pre-torunament rating, points accumulated during the tournament, and average rating of the seven opponents faced during the tournament. Tidyverse packages including dplyr and stringr are primarily used to tidy the dataset. The raw data can be located at the GitHub repository and was loaded using version control in an R Studio Project.   

# Load Packages
```{r, message=FALSE}
library(readr)
library(tidyverse)
```

# Load tournament results text file
The file can be loaded with with the read.table function 

```{r}
# Load the file directly from GitHub repository and remove unneeded dashes
data.in <- read.table('tournamentinfo.txt', header = FALSE, sep = ",") %>% unlist() %>% str_remove_all('-')

#remove the header lines (1-4) and take a look at the structure of the object
raw.data <- data.in[-c(1:4)]
head(raw.data)
str(raw.data)
```


# Loop through the raw.data object 
There are 192 rows, 2 strings for each of the 64 players plus a blank row in between.  To store the information that is requested, a for loop will be applied to the raw data object.  For each player the requested information will be extracted and stored in a blank table.  
```{r, warning=FALSE}
#create a blank table that will contain the information 
table.store <- matrix(nrow=64, ncol=6, NA) %>% as_tibble() 
colnames(table.store) <- c('PlayerID', 'PlayerName', 'State','Points', 'PreRating','OpponentRating')

for (i in 1:64){
  
  #This will select only the information regarding a single player
  player.info <- str_c(raw.data[(i*3)-2],raw.data[(i*3)-1]) %>% str_split('\\|') %>% unlist()
  
  #playerID 
  table.store[i,1] <- i
  #get name
  table.store[i,2] <- player.info[2] %>% str_trim(side='both')
  #get state
  table.store[i,3]<- player.info[11] %>% str_trim(side='both')
  #get points 
  table.store[i,4] <- player.info[3] %>% str_trim(side='both')
  #get pre-rating
  table.store[i,5] <- player.info[12] %>% str_split('[:blank:]') %>% unlist() %>% str_extract_all('\\d{3,4}') %>% unlist() %>% nth(3)
  
}
# examine the results
table.store
```

# One more loop 
The only requested information that is missing is averge rating of each players opponents.
To calculate this average, a table is made by merging each players opponents using the opponent ID field.  
```{r}
 
for (i in 1:64){
  
  #gather information for each player
  player.info <- str_c(raw.data[(i*3)-2], raw.data[(i*3)-1]) %>% str_split('\\|') %>% unlist()
  
  #extract each of the opponentIDs from each round of the tournamanet  
  opp.id <-  player.info[4:10] %>% str_extract_all('\\d{1,2}') %>% unlist()
  
  #creat a table to store the ID and calculated average
  opp.table <- matrix(nrow=length(opp.id), ncol=2, NA) %>% as_tibble

  #name the table
  names(opp.table) <- c('PlayerID','OppScore')

  #change data type to numeric so that the join will work
  opp.table$PlayerID <- as.numeric(opp.id)
  
  #join opponents pre rating to opponent id 
  opp.table$OppScore <- left_join(opp.table, table.store, by = 'PlayerID') %>%           select(PreRating) %>% unlist() %>% as.numeric() 

  #find the mean of OppScore and write to the main storage table 
  table.store[i,6] <- mean(opp.table$OppScore) %>% round(0)
}
table.store
```

#Export as a csv
The table is now normalized and can be written to GitHub as a .csv
```{r}
write.csv(table.store, "TournamentResults.csv" )
``` 


#Award Ceremony
Based on the results, which player had the strongest tournament?  This should be considered the player that tallied the most points while facing the strongest competition.

```{r}
table.store %>% filter(Points==max(Points) & OpponentRating ==max(OpponentRating)) 
```

