---
title: "DATA607 - Project 1"
author: "Rob Welk"
date: "February 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo =FALSE)
```

#Problem Statement
A document contianing information from a chess tournament is provided as an unstructured text file. The code below attempts to tidy the string data into a normalized table that is exported as a comma separated value table.  The information to extract for each chess player includes name, state, pre torunament rating, points accumulated during the tournament, and average rating of the seven opponants faced during the tournament. The raw data can be located at the GitHub repository.   

# Load Packages
```{r, echo=F}
library(readr)
library(tidyverse)
```

# Load tournament results text file
The file can be loaded with with the read.table function as demonstrated below once the 'tournament.txt' file is  
```{r}
setwd("C:/Users/robbj/CUNY SPS/DATA607")
data.in <- read.table('tournamentinfo.txt', header = FALSE, sep = ",") %>% unlist() %>% str_remove_all('-')

#remove the header lines (1-4) and take a look at the structure of the object
raw.data <- data.in[-c(1:4)]
head(raw.data)
str(raw.data)
```


# Loop through the raw.data object 
For each of the 64 players there are 3 strings. 
```{r}
# There are now 192 rows, 2 strings for each of the 64 players plus a blank in between.  

#create a blank table that will contain the information that is requested. 
table.store <- matrix(nrow=64, ncol=6, NA) %>% as_tibble() 
colnames(table.store) <- c('PlayerID', 'PlayerName', 'State','Points', 'PreRating','OpponentRating')

# In this loop 
for (i in 1:64){
  
  player.info <- str_c(raw.data[(i*3)-2],raw.data[(i*3)-1]) %>% str_split('\\|') %>% unlist()
  
  #playerID 
  table.store[i,1] <- i
  #get name
  table.store[i,2] <- player.info[2] %>% str_trim(side='both')
  #get state
  table.store[i,3]<- player.info[11] %>% str_trim(side='both')
  #get points 
  table.store[i,4] <- player.info[3] %>% str_trim(side='both')
  #get pre-rating
  table.store[i,5] <- player.info[12] %>% str_split('[:blank:]') %>% unlist() %>% str_extract_all('\\d{3,4}') %>% unlist() %>% nth(3)
  
}
table.store
```

# Loop again to find averge rating of each players opponents
To calculate this average, a table is made by merging each players opponents using the opponent ID field.  
```{r}
 
for (i in 1:64){
  
  #gather information for each player
  player.info <- str_c(raw.data[(i*3)-2], raw.data[(i*3)-1]) %>% str_split('\\|') %>% unlist()
  
  #extract each of the opponentIDs from each round of the tournamanet  
  opp.id <-  player.info[4:10] %>% str_extract_all('\\d{1,2}') %>% unlist()
  
  #creat a table to store the ID and calculated average
  opp.table <- matrix(nrow=length(opp.id), ncol=2, NA) %>% as_tibble

  #name the table
  names(opp.table) <- c('PlayerID','OppScore')

  #change data type to numeric so that the join will work
  opp.table$PlayerID <- as.numeric(opp.id)
  
  #join opponents pre rating to opponent id 
  opp.table$OppScore <- left_join(opp.table, table.store, by = 'PlayerID') %>%           select(PreRating) %>% unlist() %>% as.numeric() 

  #find the mean of OppScore and write to the main storage table 
  table.store[i,6] <- mean(opp.table$OppScore) %>% round(0)
}
table.store
```

#Export as a csv
The table is now normalized and can be written to GitHub as a .csv
```{r}
write.csv(table.store, "C:/Users/robbj/CUNY SPS/DATA607/ChessResults.csv" )
``` 


#Award Ceremony
Based on the results, which player had the strongest tournament?  This should be considered the player that tallied the most points while facing the strongest competition.

```{r}
table.store %>% filter(Points==max(Points) & OpponentRating ==max(OpponentRating)) 
```

